---
title: "Machine Learning"
author: "Inês L. Breda"
date: "8/10/2019"
output: 
 html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Summary:
This assignment aims to develop a model that can predict if a barbell lift is being well executed or not. The data includes sensor measurements located in the belt, the forearm, the arm and the dumbell of 6 individuals. Each of the 6 individuals performed barbell lifts in 5 different ways, 1 in a correct manner (class A), and 4 in an incorrect manner (class B, C, D and E). The model further described below classifies the barbell lifts with an accuracy of 0.99. 


# Packages and data

```{r library, include=FALSE}
library(caret); library(ggplot2); library(lubridate); library(rattle)
```

## Cleaning and treating data
```{r data}
#import data
training <- read.csv(file = "pml-training.csv",header = T, sep = ",", row.names = 1, na.strings=c("NA","#DIV/0!",""))
#format date and time
training$cvtd_timestamp <- parse_date_time(training$cvtd_timestamp , orders="dmy HMS")
#remove userID to aviod correlation between classe and user
training <- training[,-1]

#remove near zero variables
DataNZV <- nearZeroVar(training, saveMetrics=TRUE) 
training <- training[,DataNZV$nzv  == "FALSE"]

#remove variables with > 50% NA values
NAColum <- colSums(is.na(training))/nrow(training) >0.5
training <- training[,NAColum == "FALSE"]
#repeat the same transformations on the testing data
test <- read.csv(file = "pml-testing.csv",header = T, sep = ",", row.names = 1, na.strings=c("NA","#DIV/0!",""))
test$cvtd_timestamp <- parse_date_time(test$cvtd_timestamp , orders="dmy HMS")
test <- test[,-1]
test <- test[,DataNZV$nzv  == "FALSE"]
test <- test[,NAColum == "FALSE"]
#dimensions of training and test after dimensioning
training <- as.data.frame(training)
test <- as.data.frame(test)
dim(training); dim(test)
```

## Model creation and evaluation
```{r model}
#creating traing and testing data
inTrain <- createDataPartition(y = training$classe, p = 0.7, list = F)
trainingData <- training[inTrain,]
testingData <- training[-inTrain,]

set.seed(1777)
#Model 1: Rpart
mod1 <- train(classe ~., data = trainingData, method = "rpart")
fancyRpartPlot(mod1$finalModel)
pred1 <- predict(mod1,testingData)
confusionMatrix(pred1, testingData$classe)
#Model 2: Boosters gbm
mod2 <- train(classe ~., data = trainingData, method = "gbm", verbose = FALSE)
pred2 <- predict(mod2, testingData)
confusionMatrix(pred2, testingData$classe)
```

Model 2 is the best with an accuracy of 0.99.

## Quizz
```{r quizz}
predTest <- predict(mod2, test)
data.frame(test$problem_id, predTest)
```

